---
title: "Group model solution landscapes"
author: "Mikael Ohlsson"
date: "2023-11-21"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(ggpmisc)
{if(!require("ComplexHeatmap")) {
  library(devtools)
  install_github("jokergoo/ComplexHeatmap")
  require("ComplexHeatmap")}
  else require("ComplexHeatmap")}
library(cowplot)
#library(magrittr)
library(RColorBrewer)
library(grid)
library(scales)
library(NetIndices)
#library(FactoMineR)
#library(factoextra)
library(igraph)
library(network)
#library(gt)
library(ggtext)
library(ggalluvial)
library(ggrepel)
#library(lme4)
library(xtable)

theme_set(theme_bw())
theme_update(panel.grid = element_blank(),
             strip.background = element_rect(fill = "white", colour = "black"))
```

## Food web data formatting

Makes adjacency matrices for each iteration to be run by the group model, and separate files for their species data.

```{r FoodWebData}
weblist <- list.files(path = "../data/food_webs_pw") 
webshort <- str_replace(weblist, ".txt", "")

# Create folders if needed
if(length(list.dirs(path = "../data/adjmat/")) == 0) dir.create(path = "../data/adjmat")
if(length(list.dirs(path = "../data/sp/")) == 0) dir.create(path = "../data/sp")

# Check if adj matrices already exist
if(length(list.files(path = "../data/adjmat")) <500) {
# If not, get on with it
  for(n in 1:length(weblist)){
    net.adj <- read_delim(file = paste0("../data/food_webs_pw/", weblist[n]), delim = " ") %>%
    graph.data.frame() %>% 
    as_adjacency_matrix() %>% 
    as.matrix() %>%
    as.data.frame()
    
    # 100 iterations of adjacency matrices
    for(i in 1:100){
      
      write_delim(x = net.adj,
                   file = paste0("../data/adjmat/", 
                                 webshort[n],
                                 "_", 
                                 str_pad(i, width = 4, pad = "0"), 
                                 ".txt"),
                   delim = " ",
                   col_names = F)
      
      #And separate files with species names associated to specific rows/columns
      net.adj %>%
           rownames_to_column() %>%
           pull(var=1) %>%
           list() %>%
           fwrite(file=paste0("../data/sp/sp_",
                              webshort[n],
                              "_",
                              str_pad(i, width = 4, pad = "0"),
                              ".txt"))
    }
  }
}
```

# Group model runs

Running the group model for all the generated adjacency matrices. As we manually created separate adjacency matrix iterations, we set number of "runs" here to 1.

```{r groupmodel}
#First, checking if result files already exist. If not, running the group model. Very time consuming unless run in parallel.

if(length(list.files("../results/original")) < 500){
  setwd("../GroupModelAlgorithm")
  adjfolder <- "../data/adjmat/"
  
  netlist <- list.files(path = adjfolder, pattern = "*.txt$", full.names = T)
  netnames <- list.files(path = adjfolder, pattern = "*.txt$", full.names = F)
  
  #Number of MCMC steps
  steps <- as.character("200000")
  
  #Number of MCMC chains
  chains <- 20
  
  #Number of runs per network (runs with identical solutions will overwrite each other and thus cant be tallied)
  runs <- 1
  
  #Max number of groups per network
  groups <- 20 
  
  for(i in 1:length(netlist)){
    for(j in 1:runs) {
      r <- as.character(floor(runif(1, min=100001, max=1000000))) #random seed
    exec <- paste("./FindGroups", ncol(read.delim(netlist[i], sep = " ", header=F)), netlist[i], r, steps, chains, groups, 0) 
  #AFTER groups, 0 flag to deactivate degree correction  OR
  #1, 1, 1 flags to activate degree correction
  system(exec)
    }
  }
  setwd("../code")
  
  # Group model output data uses the same folder as source food webs. Hence, after finishing, moving group model result files to result folder:
  if(!"original" %in% list.dirs("../results", full.names = F)) dir.create("../results/original")
  
  movefiles <- list.files("../data/adjmat", pattern = ".txt-G")
  targetlocation <- paste0("../results/original/", movefiles)
  file.rename(from = paste0("../data/adjmat/", movefiles), to = targetlocation)
}
```

# Fetch group data function

Function to generate data frame with results from group model

```{r CollectResults, message = FALSE, warning = FALSE}
Collect_Results <- function() {
  results <- list.files(path = paste0("../results/original"), pattern = "Marginal") #%>% 
    #str_subset("_0\\.60_", negate = TRUE) %>% #Exclude 0.6 fraction
    #as_tibble() %>% 
    #separate(col = value, into = c("run", "marginal"), 
    #         sep = "Marginal", remove = F) %>% 
    #group_by(run) %>% 
    #summarise(marginal = max(marginal)) %>% #If a single iteration has multiple runs, select the grouping with highest marginal likelihood
    #ungroup() %>% 
    #transmute(run = paste0(run, "Marginal", marginal)) %>%
    #pull(run)
    #list.files(path = paste0("../results/newgroups/",diri,"/."), pattern = "Marginal") %>% 
    #str_subset("_0\\.60_", negate = TRUE)
  
  species <- list.files(path = paste0("../data/sp"), pattern = "*.txt$") 
  
  df <- tibble(web = "a", 
               #frac = 0.1, 
               #cuttype = "a", 
               iter = 0.1,
               marginal = 0.1,
               species = "a",
               group = 0.1) %>% 
    .[-1,]
  
    # list.files(path = paste0("../data/", resfolder), pattern = "*.txt-G") %>%
    # as_tibble() %>%
    # separate(value, c("SID", "def"), sep = ".txt", remove = F) %>%
    # separate(def, c("rest", "Marginal"), sep = "Marginal-")
  
  for(i in 1:length(results)) {
    tmp_grp <- scan(file = paste0("../results/original/",results[i]))
    tmp_sp <- scan(file = paste0("../data/sp/",species[i]), sep = "\n", what = "character")
    tmp_df <- tibble(web = substr(results[i], 1, 5),
                     #frac = as.numeric(substr(results[i], 7, 10)),
                     #cuttype = paste0(substr(results[i], 12, 13), substr(results[i], 23, 23)),
                     iter = as.numeric(substr(results[i], 7, 10)),
                     marginal = as.numeric(substr(results[i], 51,75)),
                     species = tmp_sp,
                     group = tmp_grp + 1
                     ) %>% 
      mutate(species = str_replace_all(species, "\\.", "_")) %>% 
      mutate(species = str_replace_all(species, " ", "_"))
    df <- bind_rows(df, tmp_df)
  }
  return(df)
}


#Potentially to be used for plot titles etc

webrename <- function(df) {
  df %>%
  mutate(web = case_when(web == "baren" ~ "Barents Sea",
            web == "kongs" ~ "Kongsfjorden",
            web == "reefs" ~ "Reef",
            web == "stmks" ~ "St. Marks",
            web == "ythan" ~ "Ythan")) %>%
    return()
}

group_df <- Collect_Results()

```


# Function for Jaccard Distance

Finds the best group structure (marginal likelihood-based) for each food web, and compares it to all other group structure iterations.

```{r JaccardDistance, message = FALSE, warning = FALSE}
Jaccard_Distance <- function(df){ 
  GroupAll <- tibble()
  webs <- df %>%
    distinct(web) %>%
    pull()

  # Create one DF per web and cut type (link removal, species removal).
  # base_df is the "best" group solution to which all other networks will be compared
  # Only link removal ("lr0") has 100 iterations of fraction 1, thus only selecting from those results
  # compare_df is the full collection of group results (i.e. all iterations for all fractions removed)
  # 
  for(w in 1:length(webs)){
    base_df <- df %>% # Base is the "best" grouping as in highest marginal likelihood
      filter(web == webs[w]) %>%
      filter(marginal == max(marginal)) %>% 
      group_by(species) %>% 
      slice(1) %>% # If multiple iterations give the same grouping, picking the first one
      ungroup() %>% 
      dplyr::select(species, group) # Best (marginal likelihood) original web group structure to which the reduced networks are compared
 
      time1 <- Sys.time()
      compare_df <- filter(df, web == webs[w])
        
      tmp_df <- compare_df %>% 
      dplyr::select(iter, species, group) %>%
      left_join(base_df, ., by=c("species")) 
        
        fetchsp.x <- function(g1) {
          base_df %>% 
            filter(group == g1) %>%
            pull(species) %>%
            return()
        }
          
        fetchsp.y <- function(i,g2) {
          tmp_df %>% 
            filter(iter == i & group.y == g2) %>%
            pull(species) %>%
            return()
        }
        
            GroupJ <- tmp_df %>% 
              group_by(iter, group.x) %>% 
              count(group.y) %>% 
              filter(n == max(n)) %>% 
              slice(1) %>%
              ungroup() %>%  
              mutate(sp.x = pmap(.l = list(group.x), fetchsp.x)) %>%
              mutate(sp.y = pmap(.l = list(iter, group.y), fetchsp.y)) %>%
              rowwise() %>%
              mutate(sp.intersect = list(intersect(unlist(sp.x), unlist(sp.y)))) %>%
              mutate(sp.union = list(union(unlist(sp.x), unlist(sp.y)))) %>% 
              mutate(GroupJs = length(unlist(sp.intersect)) / length(unlist(sp.union))) %>%
              ungroup() %>%
              mutate(web = webs[w])
            
            GroupAll <- bind_rows(GroupAll, GroupJ)
  
            time2 <- Sys.time()
            dtime <- as.numeric((time2-time1), units = "mins")
            
            message(sprintf("web %s taking %g minutes", webs[w], dtime))
  }
            return(GroupAll) 
}

# Average Jaccard similarity per web and fraction
WebJs <- function(df) {
  df %>% 
    group_by(web) %>% 
    summarise(Js = mean(GroupJs)) %>% 
    ungroup () %>%
    return()
}

# Web Jaccard similarity for each fraction and iteration
IterJs <- function(df) {
  df %>% 
    group_by(web, iter) %>% 
    summarise(Js = mean(GroupJs)) %>% 
    ungroup () %>%
    return()
}

# Jaccard Similarity for each group
GroupJs <- function(df) {
  df %>% 
    group_by(web, group.x) %>% 
    summarise(GroupJs = mean(GroupJs)) %>% 
    ungroup () %>%
    return()
}

J_df <- Jaccard_Distance(group_df)
```

## Jaccard similarity (Fig. 1)

Group structure similarity measured in Jaccard similarity for 100 iterations per food web, when compared to their respective (ML-based) best group structures. 

```{r Js_plot, fig.height=1.75, fig.width=6}
gg_Js <- J_df %>%  
  IterJs() %>% 
  webrename() %>%
  ggplot() +
  aes(x = Js) +
  geom_histogram(binwidth = 0.02, position = position_dodge()) +
  facet_grid(.~web) +
  labs(y = "Bin count", x = "Jaccard similarity")
gg_Js 
ggsave(filename = "../results/fig1.pdf", height = 1.75, width = 6)
```

Latex formatted table with the top 10 best solutions and the number of iterations which resulted in the same solutions (n)

```{r}
group_df %>% 
  distinct(web, iter, marginal) %>% 
  group_by(web) %>% 
  count(marginal) %>% 
  arrange(web, -marginal) %>% 
  left_join(group_df %>% 
              dplyr::select(web, iter, marginal) %>%
              group_by(web, marginal) %>%
              slice(1), #if multiple webs have the same ML, just pick one
            by = c("web", "marginal")) %>%
  slice(1:10) %>% # select top 10 ML per web
  dplyr::select(-iter) %>%
  #filter(web %in% c("baren","stmks","ythan")) %>%
  webrename() %>%
  xtable()

```

# Compare all combinations of group structures

Basis for the solution landscape plots. Group structure comparisons of all possible combinations of group structure iterations calculated. Takes about 10 minutes.

```{r}
AllJs <- function() {
compare_df <- group_df 

#Checking which species are present in the best fitting groups
checksp <- function(w, i, s){
  compare_df %>%
    filter(web == w & species %in% unlist(s)) %>%
    group_by(web, iter) %>%
    count(group) %>%
    filter(n == max(n)) %>% 
    group_by(web, iter) %>%
    slice(1) %>% #Pick one iteration in case of multiple iterations with the same group structure
    ungroup() %>%
    left_join(dplyr::select(compare_df, web, iter, group, species), 
              by = c("web", "iter", "group")) %>%
    group_by(web, iter, group, n) %>%
    nest() %>%
    rename(species = data) %>%
    return()
}

#Measuring Jaccard index, takes a couple of minutes (9 min?)
J_tmp <- compare_df %>% #filter(web == "stmks") %>%
  as_tibble() %>%
  group_by(web, iter, marginal, group) %>%
  nest() %>% #Nest species
  rename(sp = data) %>%
  ungroup() %>%
  mutate(sp.best = pmap(list(web, iter, sp), checksp)) %>%
  unnest(sp.best, names_sep = "_") %>%
  rowwise() %>%
  mutate(sp.union = list(union(unlist(sp), unlist(sp.best_species)))) %>%
  mutate(J_group = sp.best_n / length(sp.union)) %>%
  ungroup() %>%
  rename(ix = iter, iy = sp.best_iter,
          mx = marginal, Js = J_group)
}

AllJs_df <- AllJs() 

```

# Solution landscapes (Fig. 2)

Excluding St. Marks and Ythan due to very low variability. The Barents Sea food web is highly concentrated to two solutions and thus does not display a density plot.

```{r}
dist_plot <- function(w) {
  
  #For fancy plot titles
  fancy_w <- tibble(web=w) %>% webrename() %>% pull(web)  
    
  # Summarising the groupwise jaccard indices to iterations and naming iterations according to fraction links removed
  dist_df <- AllJs_df %>% ungroup() %>%
    group_by(web, ix, mx, iy) %>% 
    summarise(Js = mean(Js)) %>% 
    ungroup() %>% 
    filter(web == w) #%>% 
    # mutate(A_f_i = paste0("","_",ix),
    #        B_f_i = paste0("","_",iy))
  
  #Convert to matrix
  dist_mx <- dist_df %>% 
    #filter(sp.best_frac == 1) %>% # # # # # #
    dplyr::select(ix, iy, Js) %>%
    pivot_wider(names_from = "ix", values_from = "Js") %>%
    dplyr::select(-iy) %>%
    as.matrix()
    
  # What is the matrix ... row names
  dimnames(dist_mx)[1] <- dist_df %>% #filter(sp.best_frac == 1) %>% # ## # #
    dplyr::select(ix, iy, Js) %>%
    pivot_wider(names_from = "ix", values_from = "Js") %>% 
    pull(iy) %>% 
    list()
    
  # Run a PCA to reduce dimensions
  pcatmp <- dist_mx %>% 
    PCA(graph = F)
  
  # FactoMineR PCA plot to get coordinates
  gtmp <- pcatmp %>%
    plot(choix = "ind")
    
  # Grab data from the previous plot and add fractions and marginal likelihoods to the iterations
  ggdata <- ggplot_build(gtmp)$data[[4]] %>%
    left_join(dist_df %>% 
                dplyr::select(ix, mx) %>%
                distinct() %>%
                mutate(ix = as.character(ix)), by = c("label" = "ix")) %>%
    rowwise() %>% 
    #Adding a (visually non-noticeable) noise to help plotting really dense areas
    mutate(x = x * (1-0.001*runif(1,-1,1)), 
           y = y * (1-0.001*runif(1,-1,1))) %>%
    ungroup()
  
  # Used to expand plot area slightly to avoid clipping points at the edges
  e_x <- 0.05*diff(c(min(ggdata$x),max(ggdata$x)))
  e_y <- 0.05*diff(c(min(ggdata$y),max(ggdata$y)))
  
  # Make a complete ggplot of the PCA
  ggdata %>%
    as_tibble() %>%
    mutate(label = str_replace(label, "1_", "")) %>%
    ggplot() +
    aes(x,y, label = label, color = mx) +
    stat_density2d_filled(aes(alpha = ..level..), fill = "skyblue", # 
                          bins = 5, adjust = 5/6) +
    geom_jitter(size = 3, show.legend = F) + 
    scale_alpha_discrete(range = c(0.25,1), labels = c("low","","","","high")) +
    # geom_text_repel(max.overlaps = 200, size = 3, alpha = 1/2, fontface = "bold") +
    scale_color_distiller(palette = "RdYlGn", direction = 1) +
    #scale_fill_distiller(palette = "Spectral", direction = -1) +
    #scale_x_continuous(exp)
    scale_x_continuous(limits = c(min(ggdata$x)-e_x, max(ggdata$x)+e_x)) +
    scale_y_continuous(limits = c(min(ggdata$y)-e_y, max(ggdata$y)+e_y)) +
    guides(alpha = guide_legend(order = 1), fill = guide_legend(order = 2)) +
#    scale_fill_gradient(low = "lightblue", high = "#7777BB") +
    labs(title = fancy_w,#paste0("PCA of the ", fancy_w, " group structure iteration dissimilarities"), 
         #subtitle = "Group structure dissimilarity (point distance) and solution marginal likelihood (color)",
         color = "Marginal\nlikelihood",
         alpha = "Point\ndensity",
         shape = "Fraction links") + 
    coord_cartesian(expand = FALSE) +
#    lims(x = c(-46,15), y = c(-10,12)) +
    labs(x = paste0("Dim1 (",round(pcatmp$eig[1,2],1), "% explained)" ),
         y = paste0("Dim2 (",round(pcatmp$eig[2,2],1), "% explained)" )) + 
    theme_bw() +
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white")) %>%
    return()
}

dist_plot("baren") # Not enough spread for density plot
dist_plot("kongs")
dist_plot("reefs")
#dist_plot("stmks") # very limited variation
#dist_plot("ythan") # no variation (mind, only jitter variation if still plotting)
plot_grid(dist_plot("kongs"), dist_plot("reefs"), ncol = 2) %>%
ggsave(filename = "../results/fig2.pdf", height = 4, width = 9.75)


```


# Pairwise species stability analysis

Here we look at how often species pairs are in the same group in different group structure iterations of the same network.

```{r GroupStabFunc_both}
GroupStability <- function(webname) {
  StabDF <- tibble(Var1="a", Var2="a", value=0.1, web="a") %>% .[-1,]
    
    # The number of times the species pairs are in the same group
    for(w in 1:length(webname)){
        group_matrix <- group_df %>% 
          filter(web == webname[w]) %>%
          dplyr::select(species, group, iter) %>% 
          mutate(n = 1) %>% 
          spread(species, n, fill = 0) %>%
          dplyr::select(-group, -iter) %>%
          {crossprod(as.matrix(.))}
        
      # The number of times they occur in the food web (i.e. 100%)
        total_matrix <- group_df %>% 
          filter(web == webname[w]) %>%
          dplyr::select(species, iter) %>% 
          mutate(n = 1) %>% 
          spread(species, n, fill = 0) %>%
          dplyr::select(-iter) %>%
          {crossprod(as.matrix(.))}
        
        # The proportion of which the species pairs are grouped together
        StabTmp <- (group_matrix / total_matrix) %>%
          as.data.frame.table() %>% as_tibble()
        
        StabDF <- bind_rows(StabDF, tibble(Var1 = StabTmp$Var1, 
               Var2 = StabTmp$Var2, 
               value = StabTmp$Freq, 
               web = webname[w]))
  }
  return(StabDF)
}
```

## Running group stability function

Also adds original group membership to species from the best fit grouping (ML-based).

```{r RunGroupStab, message = FALSE, warning = FALSE}
basicwebs <- c("baren", "kongs", "reefs", "stmks", "ythan")

StabDF <- GroupStability(basicwebs)

StabDF.addgroups <- function(df) {
  webname <- StabDF %>% distinct(web) %>% pull()
  StabDF_gr <- list()

  for(i in 1:length(webname)){
    StabDF_gr[[i]] <- StabDF %>% 
      filter(web == webname[i]) %>%
      left_join(group_df %>% 
                  filter(web == webname[i]) %>% 
                  filter(marginal == max(marginal)) %>%
                  group_by(species) %>%
                  slice(1) %>%
                  dplyr::select(species, web, group), 
                by = c("Var1" = "species", "web")) %>%
      left_join(group_df %>% 
                  filter(web == webname[i]) %>% 
                  filter(marginal == max(marginal)) %>%
                  group_by(species) %>%
                  slice(1) %>%
                  dplyr::select(species, web, group), 
                by = c("Var2" = "species", "web"))
  }
  StabDF_gr <- tibble(A = StabDF_gr) %>% unnest(cols = c(A))
  return(StabDF_gr)
}
StabDF_gr <- StabDF.addgroups(StabDF)
```

### Heatmap function

Arranges species for each food web along the x- and y-axis according to their best group structures.

```{r ComplexHeatmap, fig.width=15, fig.height=9, out.width="100%", message=FALSE, warning=FALSE}

# Function used for plotting purposes when arranging species.
original_groups <- function(webname) {
  original <- group_df %>% 
    filter(web %in% webname) %>%
    filter(marginal == max(marginal)) %>%
                  group_by(species) %>%
                  slice(1) %>%
    ungroup() %>%
    dplyr::select(species, group, iter) 
  sp_order <- original %>% arrange(group)# %>% pull(species)
  return(sp_order)
}


# Main plotting function

CheatPlot <- function(webname) {
  # Checking original number of groups (used for coloring)
  webtitle <- tibble(web = webname) %>% webrename() %>% pull(web)
  clen <- original_groups(webname) %>% pull(group) %>% n_distinct()

      # Heatmap Annotations with group color codings taken from the RColorBrewer package. 
   top_ha <- HeatmapAnnotation(Group = original_groups(webname) %>% arrange(group) %>%
                                 pull(group),
                           col = list(Group = tibble(A = original_groups(webname) %>% 
                                                       distinct(group) %>%
                                                       mutate(group = as.factor(group)) %>%
                                                       pull(group),
                                                     B = c(brewer.pal(name="Paired", n = 8),
                                                           brewer.pal(name="Dark2", n = 8))[1:clen]) %>%
                                        arrange(A) %>%
                                        deframe()),
                          show_legend = F,
                          show_annotation_name = F)
   
   
  
   left_ha <- rowAnnotation(Group = original_groups(webname) %>%
                              mutate(group = as.factor(group)) %>%
                              dplyr::pull(group),
                           col = list(Group = tibble(A = original_groups(webname) %>% 
                                                       distinct(group) %>%
                                                       mutate(group = as.factor(group)) %>%
                                                       pull(group),
                                                     B = c(brewer.pal(name="Paired", n = 8),
                                                           brewer.pal(name="Dark2", n = 8))[1:clen]) %>%
                                        arrange(A) %>%
                                        deframe()),
                          show_annotation_name = F)
    
    pheatdata <- StabDF_gr %>% 
      dplyr::select("Var1", "Var2", "value", "web") %>% 
      filter(web == webname) %>%
      pivot_wider(values_from = value, names_from = c(Var2)) %>% 
      dplyr::select(-c("web")) %>%
      set_rownames(.$Var1) %>% 
      data.matrix(rownames.force = T) %>% 
      .[,-1] %>% 
      .[(original_groups(webname) %>% arrange(group) %>% dplyr::pull("species")),] %>%
      .[,(original_groups(webname) %>% arrange(group) %>% dplyr::pull("species"))]
    
    Heatresult <- ComplexHeatmap::Heatmap(pheatdata, name = "Same \ngroup",
                            col = colorRampPalette(c("white", "grey", "black"))(50),
                            cluster_rows = F,
                            cluster_columns = F,
                            show_column_names = F,
                            show_row_names = F,
                            top_annotation = top_ha,
                            left_annotation = left_ha,
                            column_title = webtitle) 
    
  return(Heatresult)
}
```

### Heatmaps (Fig. 3)

Species arranged in groups based on the best fit (ML) result

```{r Cheat_grid_NOR, fig.width=15, fig.height=14, out.width="100%", warning=FALSE}
cheat_baren <- CheatPlot("baren")
cheat_kongs <- CheatPlot("kongs")
cheat_reefs <- CheatPlot("reefs")
cheat_stmks <- CheatPlot("stmks")
#cheat_ythan <- CheatPlot("ythan") # Skipping Ythan as it only has one group structure solution

cheat_grid <- plot_grid(grid.grabExpr(draw(cheat_baren)),
          grid.grabExpr(draw(cheat_kongs)), grid.grabExpr(draw(cheat_reefs)),
          grid.grabExpr(draw(cheat_stmks)), #grid.grabExpr(draw(cheat_ythan)),
          ncol = 2, 
          #labels = c("A", "B", "C", "D", "E"),
          hjust = -0.4)

cheat_grid
ggsave(filename = "../results/fig3.pdf", plot = cheat_grid, width = 9.5, height = 9)


```

# Species data

Species taxonomies for the Barents Sea were obtained from Kortsch et al. (2018), while taxonomies for all other networks were obtained from  Cirtwill & Eklöf (2018, and references therein).

```{r}
TraitsDF <- read_csv(file = "../data/food_webs_data/full_data_upd.csv") %>%
  mutate(Network = case_when(Network == "baren" ~ "Barent's Sea",
            Network == "baske" ~ "Serengeti",
            Network == "kongsfjorden" ~ "kongs",
            Network == "reef" ~ "reefs",
            Network == "stmarks" ~ "stmks",
            Network == "ythanjacob" ~ "ythan")) %>%
  dplyr::select(Network, Species, Group, Kingdom, Phylum, Class, Order, Family, 
                Genus, BodyWeight, FeedingMode, MetabolicCategory, FeedingType, 
                Mobility, Environment) %>%
  bind_rows((read_tsv(file = "../data/kortsch/SpeciesList.txt") %>% #Limited to taxonomy for Barents Sea, also adding genus column from species name
  rename(Species = ABBREVIATION, Phylum = PHYLUM_SUBPYLUM, 
         Class = CLASS, Order = ORDER, Family = FAMILY) %>% 
  mutate(Network = "baren") %>% 
  mutate(TROPHOSPECIES = str_to_sentence(TROPHOSPECIES)) %>%
  separate(TROPHOSPECIES, into = c("Genus", "xx"), sep = "_") %>%
  dplyr::select(-"xx", -"GROUP"))
              ) %>%
      mutate(Species = str_replace_all(Species, "\\.", "_")) %>%
      mutate(Species = str_replace_all(Species, " ", "_"))

StabDF_Traits <- StabDF_gr %>% 
  mutate(Var1 = str_replace(Var1, pattern="___", replacement="_"),
         Var2 = str_replace(Var2, pattern="___", replacement="_")) %>%
  left_join(TraitsDF, by = c("Var1" = "Species", "web" = "Network")) %>%
  left_join(TraitsDF, by = c("Var2" = "Species", "web" = "Network"))

```

## Taxonomic similarities (Fig. 4)

Adding closest taxonomic relationship for all species pairs which were found in the same group.

```{r Taxonomy_simi, fig.width=10, fig.height=5, out.width="100%"}
gg_pw_close_taxa <- StabDF_Traits %>%
  filter(Var1 != Var2) %>%
  filter(group.x == group.y & web %in% c("baren", "kongs", "reefs")) %>%
  mutate(TaxSim = ifelse(!is.na(Phylum.x) & !is.na(Phylum.y) & Phylum.x == Phylum.y, "Phylum", "Unrelated")) %>%
  mutate(TaxSim = ifelse(!is.na(Class.x) & !is.na(Class.y) & Class.x == Class.y, "Class", TaxSim)) %>%
  mutate(TaxSim = ifelse(!is.na(Order.x) & !is.na(Order.y) & Order.x == Order.y, "Order", TaxSim)) %>%
  mutate(TaxSim = ifelse(!is.na(Family.x) & !is.na(Family.y) & Family.x == Family.y, "Family", TaxSim)) %>%
  mutate(TaxSim = ifelse(!is.na(Genus.x) & !is.na(Genus.y) & Genus.x == Genus.y, "Genus", TaxSim)) %>%
  mutate(TaxSim = as.factor(TaxSim),
    TaxSim = fct_relevel(TaxSim, c("Unrelated", "Phylum", "Class", "Order", "Family", "Genus"))) %>%
    group_by(grp = paste(pmax(Var1, Var2), pmin(Var1, Var2), sep = "_")) %>%
  distinct(grp, .keep_all = TRUE) %>%
  ungroup() %>%
  dplyr::select(-grp) %>%
  webrename() %>%
  ggplot() +
  aes(x = TaxSim, y = value, fill = TaxSim, group = TaxSim) + 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  facet_grid(.~web, scales = "free_x", space = "free") +
  scale_y_continuous(expand = c(0,0.03), limits = c(0,1)) +
  scale_fill_manual(values = c("#555555",rev(brewer.pal(5, "RdYlBu")))) +
  labs(x = "Species pairs' closest taxonomic relationship", y = "Pairwise stability") +
  theme(axis.text.x = element_text(angle = -30, hjust=0))
gg_pw_close_taxa
ggsave(filename = "../results/fig4.pdf", plot = gg_pw_close_taxa, width = 7, height = 3)
```

# Jaccard similarity vs Marginal likelihood (Fig. S1)

```{r}
Js_ML <- group_df %>% 
  filter(!web %in% c("ythan")) %>% 
  dplyr::select(web, iter, marginal, group) %>%
  distinct() %>%
  left_join(J_df %>%
    dplyr::select(web, iter, group.x, GroupJs), 
    by = c("web", "iter", "group" = "group.x")) %>% 
  group_by(web, iter) %>%
  summarise(marginal = mean(marginal), Js = mean(GroupJs, na.rm = T)) %>% 
  ungroup() %>%
  distinct(web, marginal, Js) %>%
  webrename() 

ML_borders <- Js_ML %>% group_by(web) %>% mutate(ML_3 = max(marginal) - 3,
                                   ML = max(marginal),
                                   ML_min = min(marginal),
                                   JS_min = min(Js)) %>%
  ungroup() %>% 
  distinct(web, ML_3, ML, ML_min, JS_min)

 
gg_Js_ML <- ggplot() + 
  geom_point(data = Js_ML, aes(x = marginal, y = Js)) + 
  facet_wrap(web~., scales = "free") +
  labs(x = "Marginal likelihood", y = "Jaccard similarity (vs. best solution)")
gg_Js_ML
ggsave(filename = "../results/figs1.pdf", plot = gg_Js_ML, width = 6.5, height = 6)

```

# Alluvial plot (Fig. S2)

An example plot visualizing the structural differences between two group structure iterations in Kongsfjorden.

```{r alluvials}

AlluPlotter <- function(u,v, net) {
AlluDF <- J_df %>%
  IterJs() %>%
  filter(web == net & 
           iter %in% c(u,v)) %>%
  mutate(.r = ifelse(iter == u, 1, 2)) %>%
  left_join(group_df, by=c("web", "iter")) %>% 
  left_join(TraitsDF, by = c("web" = "Network", "species" = "Species")) 

  # Can be used for plot title
  CurrentJs <- AllJs_df %>% 
    filter(web == net, ix == u, iy == v) %>%
    pull(Js) %>%
    mean() %>%
    round(3)
  
  # Can be used for plot title
  CurrentWeb <- AllJs_df %>% 
    filter(web == net, ix == u, iy == v) %>%
    webrename() %>% 
    pull(web) 

  ggplot(AlluDF, aes(x=as.factor(.r), stratum=as.factor(group),
                                   alluvium=species, fill=as.factor(Phylum),
                                   label = as.factor(group))) +
  geom_flow(stat="alluvium", position="identity", reverse=F, lode.guidance="rightward", alpha = 1) + #aes.flow="forward", aes.bind=T,
  geom_stratum(reverse=F, width = 1/3, fill = "grey") +
  geom_text_repel(stat = "stratum", reverse=F, force = 0.0002) +
  theme_classic() +
  scale_x_discrete(labels = c(u,v), expand=c(0.2,0)) +
  scale_y_discrete(labels=AlluDF$species, name="Species") +
  scale_fill_manual(name="Phylum", values = c(brewer.pal(name="Set1", n = 9),
                                             brewer.pal(name="Set2", n = 8),
                                             brewer.pal(name="Paired", n = 12)
                                             )[1:29]) + 
  theme(axis.text.x = element_text(size=12),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.y = element_blank()) +
  guides(fill = guide_legend(ncol = 1)) +
  labs(#title = paste0("Original variation of ", CurrentWeb), 
       #subtitle = paste0("Displaying iterations ", u, " and ", v, " with Jaccard similarity ", CurrentJs),
       x = "Iterations") %>%
  return()
}
gg_KongsAllu <- AlluPlotter(63,47,"kongs") #Js 0.856
gg_KongsAllu
ggsave(filename = "../results/figs2.pdf", height = 7, width = 5)
```
